---
title: "determining_fire_ephemerals"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


standard library load
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(png)
library(raster)
library(rgdal)
library(ggplot2)
library(knitr)
library(leaflet)
library(maps)
library(lubridate)
library(RColorBrewer)
library(tiff)
library(plyr)
```





Fire data:
   can be downloaded from here - (url)
   It is modis data seperated into each month for each year from nov 2000 - 2021
   Other fire data could be used with code modifications. 

load in function
```{r}
burn_data <- function(path, species_df) {
# Gathering all the file names
   all_numbers<-str_extract(path, "(?<=A)[0-9]*")
#Getting dates from file names
   year<-substr(all_numbers, start = 1, stop = 4)
   julian_date<-substr(all_numbers, start = 5, stop = 7)
#Extracting fire dates for each observation location  
    fire_raster <- raster::stack(path)
  species_df$burn_data_julian_date <-
    raster::extract(fire_raster,
            cbind(species_df$decimalLongitude, species_df$decimalLatitude))
#Formating fire data dates  
  species_df$fire_year <- year
  fire_date <- as.Date(paste0(year, "-01-01"))
  yday(fire_date) <- as.numeric(julian_date)
  species_df$fire_date <- fire_date
  species_df$fire_month <- month(fire_date)
  return(species_df)
}
```


Species data:
  As an example we use Actinotus forsythii data was downloaded from ALA
you will need species data with - latitude and longditude location and date of record (day, month and year). It also useful to have location description. 

```{r}
Actinotus_forsythii_alaobvs_28_10_21 <- read_csv("Actinotus_forsythii_alaobvs_28_10_21.csv")

Actinotus_forsythii <- dplyr::select(Actinotus_forsythii_alaobvs_28_10_21, raw_recordedBy, basisOfRecord, year, month, day, verbatimLocality, decimalLatitude, decimalLongitude)
```

Running the load in function with apropriate species data
```{r}
#List of file names
list_of_files <- list.files("Win20", recursive =  TRUE, full.names = TRUE) %>% str_subset("burndate")

#Large dataframe with all months/years at each observation location (fire and non-fire results)
#Species_df can be changed to whichever species is being used
output<-lapply(list_of_files, burn_data, species_df=Actinotus_forsythii)
big_df<-bind_rows(output)

#Subsetting for records where fire was present (and including location observation in the relervent timeframe)
fire_only_records <- filter(big_df, burn_data_julian_date!=0 & year>=2000,year > fire_year)
fire_only_records$collection_date <- ymd(paste(fire_only_records$year,fire_only_records$month,fire_only_records$day,sep="-")) 
fire_only_records$date_differnce<-as.numeric(fire_only_records$collection_date-fire_only_records$fire_date)

```

calculations
```{r}
hist(fire_only_records$date_differnce)

##2 years of fire history
fire_only_records %>% filter(date_differnce<730)->recent_fires

fire <- nrow(recent_fires)
total <- nrow(Actinotus_forsythii)
percentage_burnt <- (fire/total)*100

```

My A. forsythii data came back 37% fire present in the past two years













running same code for Swainsona formosa

```{r}
Swainsona_formosa_records <- read_csv("Swainsona formosa_records-2021-11-24.csv")

Swainsona_formosa <- dplyr::select(Swainsona_formosa_records, basisOfRecord, year, month, day, decimalLatitude, decimalLongitude)
```

```{r}
#Large dataframe with all months/years at each observation location (fire and non-fire results)
#Species_df can be changed to whichever species is being used
output<-lapply(list_of_files, burn_data, species_df=Swainsona_formosa)
big_df<-bind_rows(output)

#Subsetting for records where fire was present (and including location observation in the relervent timeframe)
fire_only_records <- filter(big_df, burn_data_julian_date!=0 & year>=2000,year > fire_year)
fire_only_records$collection_date <- ymd(paste(fire_only_records$year,fire_only_records$month,fire_only_records$day,sep="-")) 
fire_only_records$date_differnce<-as.numeric(fire_only_records$collection_date-fire_only_records$fire_date)

```

calculations
```{r}
##2 years of fire history
fire_only_records %>% filter(date_differnce<730)->recent_fires

fire <- nrow(recent_fires)
total <- nrow(Swainsona_formosa)
percentage_burnt <- (fire/total)*100


hist(fire_only_records$date_differnce)
ggsave("Swainsona_formosa_date_frequency.jpg")

```

S. formosa came back 20% fire in the past 2 years however the overall histagram looks very different



Extra notes
```{r}
Swainsona_formosa_fire_only_records <- fire_only_records
Swainsona_formosa_recent_fires <- recent_fires

```


