---
title: "WIP_code"
output: html_document
---

```{r}
install.packages("knitr")
library(knitr)

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("plyr")
```






standard library load
```{r}
library(readr)
library(plyr)
library(dplyr)
library(tidyverse)
library(png)
library(raster)
library(rgdal)
library(ggplot2)
library(knitr)
library(leaflet)
library(maps)
library(lubridate)
library(RColorBrewer)
library(tiff)
```


Actinotus forsythii data was downloaded from ALA
(this could also be turned into code that downlads straight from ala -although not particularly useful for non australian species)

```{r}
Actinotus_forsythii_alaobvs_28_10_21 <- read_csv("Actinotus_forsythii_alaobvs_28_10_21.csv")
View(Actinotus_forsythii_alaobvs_28_10_21)

Actinotus_forsythii <- dplyr::select(Actinotus_forsythii_alaobvs_28_10_21, raw_recordedBy, basisOfRecord, year, month, day, verbatimLocality, decimalLatitude, decimalLongitude)

Actinotus_forsythii <- dplyr::mutate(Actinotus_forsythii, id = row_number())

View(Actinotus_forsythii)
```



checking if lat long line up with what is expexted

```{r}

world <- map_data("world")

ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "white", fill = "lightgray", size = 0.1
    ) +
   theme_bw() +
  geom_point(
    data = Actinotus_forsythii,
    aes(decimalLongitude, decimalLatitude),
    alpha = 0.4, size = 0.1
    ) + scale_x_continuous(limits = c(140,155)) + scale_y_continuous(limits = c(-40,-30))

```


next...

Fire data needs to be attained

Fire data will be a grid where as lat, long of species data is a point. not sure if this will cause issues. 

Each location point has a date stamp. This date stap needs to be turned into a range and converted to the same format as the fire data date stamp. 

True/false for each data point if fire occours in the agreed date range. 

Percentage true false?



Using fire data from previous project (loaded into workspace from that project)
preliminalrily shows that graphically where we see fire, the following year we see A. forsythii now to try and test this.


```{r}
fire_2003_true <- subset(fire_2003, layer > 0)

Actinotus_forsythii_2003_burn <- subset(Actinotus_forsythii, year >2003 & year <2008)
```


```{r}
ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "black", fill = "white", size = 0.1
    ) +
   theme_bw() +
  geom_point(
    data = fire_2003_true,
    aes(x, y, colour=layer),
    alpha = 0.5, size = 1
    ) + scale_x_continuous(limits = c(147.5,152.5)) + scale_y_continuous(limits = c(-40,-32.5)) + 
  geom_point(
    data = Actinotus_forsythii_2003_burn,
    aes(decimalLongitude, decimalLatitude),
   alpha =0.5, size = 1
    ) + scale_x_continuous(limits = c(147.5,152.5)) + scale_y_continuous(limits = c(-40,-32.5))
```



firstly how do inflate fire data to be a range rather than a point?

```{r}

fire_2003_test <- fire_2003 


xrange <- function(fire_data) {
  fire_data %>%  mutate(xrange = x + 0.250) 
}
yrange <- function(fire_data) {
  fire_data %>% mutate(yrange = y - 0.250)
}
fire_true <- function(fire_data) {
  fire_data %>% mutate(fire_true = if_else (layer !=	0, TRUE, FALSE))
}


fire_2003_test <- xrange(fire_2003)
fire_2003_test <- yrange(fire_2003_test)
fire_2003_test <- fire_true(fire_2003_test)


str(fire_2003_test)

```

ok, now I need to crass reference. 
does point plant fit between x-xrange 


IGNORE THIS for a minute (code not runnin)
give me fire_true where (decimallat is >yrange and <y) and (decimallong >xrange and <x)

```{r}
fire_true <- (Actinotus_forsythii_2003_burn$decimalLongitude > fire_2003_test$yrange 
  & < fire_2003_test$y) &
  (Actinotus_forsythii_2003_burn$decimalLatitude > fire_2003_test$xrange 
  & < fire_2003_test$x)


get_true <-function()
  
  
fire_2003_test %<% left_join(Actinotus_forsythii_2003_burn, by = c())  

```



```{r}

burn_data <- function(path, species_df) {
   all_numbers<-str_extract(path, "(?<=A)[0-9]*")
   year<-substr(all_numbers, start = 1, stop = 4)
   julian_date<-substr(all_numbers, start = 5, stop = 7)

    fire_raster <- raster::stack(path)
  species_df$burn_data_julian_date <-
    raster::extract(fire_raster,
            cbind(species_df$decimalLongitude, species_df$decimalLatitude), buffer=100, fun=max)
  species_df$fire_year <- year
  fire_date <- as.Date(paste0(year, "-01-01"))
  yday(fire_date) <- as.numeric(julian_date)
  species_df$fire_date <- fire_date
  species_df$fire_month <- month(fire_date)
  return(species_df)
}


list_of_files <- list.files("Win20", recursive =  TRUE, full.names = TRUE) %>% str_subset("burndate")


output<-lapply(list_of_files,burn_data,species_df=Actinotus_forsythii)
big_df<-bind_rows(output)

fire_only_records <- filter(big_df,burn_data_julian_date!=0 & year>=2000,year > fire_year)
fire_only_records$collection_date <- ymd(paste(fire_only_records$year,fire_only_records$month,fire_only_records$day,sep="-")) 
fire_only_records$date_differnce<-as.numeric(fire_only_records$collection_date-fire_only_records$fire_date)


fire_only_records %>% filter(date_differnce<600)->recent_fires

fire <- nrow(recent_fires)
total <- nrow(Actinotus_forsythii)
percentage_burnt <- (fire/total)*100

```



dethermoning_fire_ephemeral Maps!!

```{r}


world <- map_data("world")

ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "black", fill = "white", size = 0.1
    ) +
   theme_bw() +
  geom_point(
    data = Actinotus_forsythii,
    aes(decimalLongitude, decimalLatitude), size = 0.1) + 
  scale_x_continuous(limits = c(147.5,152.5)) + scale_y_continuous(limits = c(-37.5,-32.5)) +
  geom_point(
    data = recent_fires,
    aes(decimalLongitude, decimalLatitude, colour= fire_year),
   size = 0.1
    ) + scale_x_continuous(limits = c(147.5,152.5)) + scale_y_continuous(limits = c(-37.5,-32.5))


```




```{r}
library(tidyverse)
library(raster)
library(lubridate)
library(parallel)
#donkey <- read_csv("observations-201459.csv")
#ducks <- read_csv("observations-201466.csv")


#files of species records from gbif
list_of_gbif_files <-
  list.files("data/gbif", recursive =  TRUE, full.names = TRUE) %>% str_subset("210914110416597")

#modis fira reccords
list_of_files <-
  list.files("Win20", recursive =  TRUE, full.names = TRUE) %>% str_subset("burndate")


burn_data_gbif <- function(path, species_df, buffer_size = 1000) {
  require(sf)
  require(lubridate)
  species_df <- dplyr::filter(species_df, !is.na(decimalLatitude)) #filtering out records with no location
  species_df <- dplyr::filter(species_df, year >= 2003) #filtering for reccords 2003 and later (why not 2001?)
    all_numbers <- str_extract(path, "(?<=A)[0-9]*") 
    year <- substr(all_numbers, start = 1, stop = 4)
    julian_date <- substr(all_numbers, start = 5, stop = 7) # getting date from the file name
  sp <-
    SpatialPoints(cbind(species_df$decimalLongitude, species_df$decimalLatitude))
  fire_raster <- raster::raster(path) 
  clip1 <-
    crop(fire_raster, extent(sp)) #this speeds things up a lot
  species_df$burn_data_julian_date <-
    raster::extract(clip1,
                  sp, buffer = buffer_size, fun = max) #extracting required information from the raster according to the species reccord points
  species_df$fire_year <- year
  fire_date <- as.Date(paste0(year, "-01-01"))
  yday(fire_date) <- as.numeric(julian_date) 
  species_df$fire_date <- fire_date
  species_df$fire_month <- month(fire_date) # changeing how the dates are displayed
  return(species_df)
}


test<-burn_data_gbif(list_of_files[[1]], Actinotus_forsythii)
#test works on one species


calculate_burn_obs <-
  function(list_of_files = list_of_files,
           species_df = species_df,
           cores = 1) {
    require(parallel)
    #hasen't this bit already been done in the burn_data_gbif function?
    species_df <- dplyr::filter(species_df, year >= 2003) %>% 
      dplyr::filter(coordinateUncertaintyInMeters < 1000 |
                      is.na(coordinateUncertaintyInMeters)) # i am unsure about the "coordinateUncertaintyInMeters"
    
    
    
output <-
      mclapply(list_of_files,
               burn_data_gbif,
               species_df = species_df,
               mc.cores = cores)
    big_df <- bind_rows(output)
    big_df$fire_date[big_df$burn_data_julian_date <= 0] <- NA
    big_df$date_differnce <-
      as.numeric(as_date(big_df$eventDate) - big_df$fire_date)
    big_df$date_differnce[big_df$date_differnce < 0] <- NA
    big_df %>%
      group_by(scientificName,
               occurrenceID,
               eventDate,
               coordinateUncertaintyInMeters,
               year) %>%
      summarize(date_difference = min(date_differnce, na.rm = T)) -> most_recent_fire_per_obs
    return(most_recent_fire_per_obs)
  }



all_gbif <-
  bind_rows(lapply(list_of_gbif_files, read_delim, delim = " "))


one <- calculate_burn_obs(list_of_files, species_df = all_gbif, cores = 4)

#Error: Problem with `filter()` input `..1`. ℹ Input `..1` is `year >= 2003`. x comparison (5) is possible only for atomic and list types Run `rlang::last_error()` to see where the error occurred.
#In addition: Warning message:
#Unknown or uninitialised column: `date_differnce`. 
```





graphs 

```{r}
one %>%
  mutate(fire = (date_difference < 1825)) %>%
  mutate(sp = word(scientificName, 1, 2)) %>%
  ggplot(aes(x = as_date(eventDate), fill = fire)) + geom_histogram() +
  facet_grid(sp ~ ., scales = "free")
ggsave("fire_yes.png", height = 11, width = 6)

one %>% dplyr::filter(date_difference < 1825) %>%
  mutate(sp = word(scientificName, 1, 2)) %>%
  ggplot(aes(x = date_difference, fill = year)) + geom_histogram() + facet_grid(sp ~
                                                                                  ., scales = "free")
ggsave("juve.png", height = 11, width = 4)

one %>%
  group_by(word(scientificName, 1, 2)) %>%
  summarise(prop_fire = sum(date_difference < 1825) / n()) %>%
  arrange(prop_fire)


one %>% dplyr::filter(date_difference < 1825) %>%
  mutate(sp = word(scientificName, 1, 2)) %>%
  mutate(date_difference_years = date_difference/365.25) %>%
  dplyr::filter(sp %in% c("Androcalva rosea","Actinotus forsythii","Caleana major","Burnettia cuneata","Blandfordia grandiflora","Actinotus helianthi")) %>%
  ggplot(aes(x = date_difference_years, fill = year)) + 
  geom_histogram() + facet_grid(sp ~ ., scales = "free")+
  theme_bw()+xlab("Time from fire to collection / observation (years)")
ggsave("selected_species.png",height=11,width=6)




```






what will wrote but for windoes

```{r}

library(tidyverse)
library(raster)
library(lubridate)
library(sf)
#donkey <- read_csv("observations-201459.csv")
#ducks <- read_csv("observations-201466.csv")

list_of_gbif_files <-
  list.files("data/gbif/", full.names = TRUE) %>% str_subset("210914110416597")



actinotus <- read_delim("data/gbif/actinotus/0081783-210914110416597.csv")


all <-
  bind_rows(lapply(list_of_gbif_files,read_delim))


burn_data_gbif <- function(path, species_df, buffer_size = 1000) {
  require(sf)
  require(lubridate)
  species_df <- dplyr::filter(species_df, !is.na(decimalLatitude))
  species_df <- dplyr::filter(species_df, year >= 2003)
  all_numbers <- str_extract(path, "(?<=A)[0-9]*")
  year <- substr(all_numbers, start = 1, stop = 4)
  julian_date <- substr(all_numbers, start = 5, stop = 7)
  sp <-
    SpatialPoints(cbind(species_df$decimalLongitude, species_df$decimalLatitude))
  fire_raster <- raster::raster(path)
  clip1 <-
    crop(fire_raster, extent(sp)) #this speeds things up a lot
  species_df$burn_data_julian_date <-
    raster::extract(clip1,
                    sp, buffer = buffer_size, fun = max)
  species_df$fire_year <- year
  fire_date <- as.Date(paste0(year, "-01-01"))
  yday(fire_date) <- as.numeric(julian_date)
  species_df$fire_date <- fire_date
  species_df$fire_month <- month(fire_date)
  return(species_df)
}


list_of_files <-
  list.files("Win20", recursive =  TRUE, full.names = TRUE) %>% str_subset("burndate")

#testing first function
Sys.time()
z <- burn_data_gbif(list_of_files[[1]], species_df = pinks)
Sys.time()


calculate_burn_obs <-
  function(list_of_files = list_of_files,
           species_df = species_df) {
    require(parallel)
    species_df <- dplyr::filter(species_df, year >= 2003) %>%
      dplyr::filter(coordinateUncertaintyInMeters < 1000 |
                      is.na(coordinateUncertaintyInMeters))
    output <-
      lapply(list_of_files,
               burn_data_gbif,
               species_df = species_df)
    big_df <- bind_rows(output)
    big_df$fire_date[big_df$burn_data_julian_date <= 0] <- NA
    big_df$date_differnce <-
      as.numeric(as_date(big_df$eventDate) - big_df$fire_date)
    big_df$date_differnce[big_df$date_differnce < 0] <- NA
    
    big_df %>%
      group_by(species,
               occurrenceID,
               eventDate,
               coordinateUncertaintyInMeters,
               countryCode,
               year) %>%
      summarize(date_difference = min(date_differnce, na.rm = T)) -> most_recent_fire_per_obs
    return(most_recent_fire_per_obs)
  }











androclava<-filter(all,genus=="Androcalva")
one <- calculate_burn_obs(list_of_files, species_df = actinotus)

one %>%
  mutate(fire = (date_difference < 1825)) %>%
  mutate(sp = word(species, 1, 2)) %>%
  ggplot(aes(x = as_date(eventDate), fill = fire)) + geom_histogram() +
  facet_grid(sp ~ ., scales = "free")
ggsave("fire_yes.png", height = 20, width = 4)




one %>% dplyr::filter(date_difference < 1825) %>%
  mutate(date_difference_months = date_difference/365.25*12) %>%
  mutate(sp = word(species, 1, 2)) %>%
  ggplot(aes(x = date_difference_months)) + geom_histogram() + facet_grid(sp ~ ., scales = "free")
ggsave("juve.png", height = 20, width = 4)

one %>%
  group_by(sp=word(species, 1, 2)) %>%
  summarise(prop_fire = sum(date_difference < 1825) / n(),n=n()) %>%
  arrange(prop_fire)->out













one %>% dplyr::filter(date_difference < 1825) %>%
  mutate(sp = word(species, 1, 2)) %>%
  mutate(date_difference_years = date_difference/365.25) %>%
  dplyr::filter(sp %in% c("Androcalva rosea","Androcalva reticulata","Androcalva pearnii","Androcalva loxophylla")) %>%
  ggplot(aes(x = date_difference_years)) + 
  geom_histogram() + facet_grid(sp ~ ., scales = "free")+
  theme_bw()+xlab("Time from fire to collection / observation (years)")
ggsave("selected_species.png",height=11,width=6)

```

```{r}

actinotus$species %>% 
  unique()





split(df, df$ids)




```



actinotus:
  helianthi
  gibbonsii
  forsythii
  leucocephalus
  
  
full actinotus list 
GBIF.org (17 December 2021) GBIF Occurrence Download https://doi.org/10.15468/dl.872w7u 


































fire ephemerals from the list

Actinotus leucocephalus, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.anh5dk
Alyogyne hakeifolia, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.gbjdth
Alyogyne huegelii, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.x5rak8
Grevillea scapigera, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.3k9rg7 (aus started here)
Cadonocarpus cotinifoliua, (no reccord)
Gyrostemon racemiger, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.kssrx2
Gyrostemon ramulosus, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.nsrh6c
Tersonia cyathiflora, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.v59y3p
Hydrocotyle serendipita, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.442xbr
Hydrocotyle phoenix, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.fkangr
Yakirra (Panicum) australiense (gramineae), (no record)
Goodenia incana,  GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.wgvunc
Tripterococcus brunonis, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.q3hsm4
Austrostipa compressa,  GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.9c3jyr
Austrostipa macalpinei (Poaceae), GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.g4phaq
Kennedia coccinea, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.ce3wr4
Actinotus forsythii, GBIF.org (02 December 2021) GBIF Occurrence Download  https://doi.org/10.15468/dl.2ghfvg







Other fire-ephemeral herbs: Dampiera sacculata, D. wellsiana, Goodenia coerulea, Platysace juncea, Chlaenosciadium gardneri, Stackhousia scoparia" 









this will clear your environment
```{r}
rm(list = ls())
```






