---
title: "WIP_code"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

standard library load
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(png)
library(raster)
library(rgdal)
library(ggplot2)
library(knitr)
library(leaflet)
library(maps)
```


Actinotus forsythii data was downloaded from ALA
(this could also be turned into code that downlads straight from ala -although not particularly useful for non australian species)

```{r}
Actinotus_forsythii_alaobvs_28_10_21 <- read_csv("Actinotus_forsythii_alaobvs_28_10_21.csv")
View(Actinotus_forsythii_alaobvs_28_10_21)

Actinotus_forsythii <- dplyr::select(Actinotus_forsythii_alaobvs_28_10_21, raw_recordedBy, basisOfRecord, occurrenceRemarks, eventDate, year, month, day, habitat, stateProvince, locality, verbatimLocality, decimalLatitude, decimalLongitude)

View(Actinotus_forsythii)
```


checking if lat long line up with what is expexted

```{r}

world <- map_data("world")

ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "white", fill = "lightgray", size = 0.1
    ) +
   theme_bw() +
  geom_point(
    data = Actinotus_forsythii,
    aes(decimalLongitude, decimalLatitude),
    alpha = 0.4, size = 0.1
    ) + scale_x_continuous(limits = c(140,155)) + scale_y_continuous(limits = c(-40,-30))

```


next...

Fire data needs to be attained

Fire data will be a grid where as lat, long of species data is a point. not sure if this will cause issues. 

Each location point has a date stamp. This date stap needs to be turned into a range and converted to the same format as the fire data date stamp. 

True/false for each data point if fire occours in the agreed date range. 

Percentage true false?


































